6.867 Problem Set 1

### 2015-10-02 pset1/paper.tex
done

>>>>

\section{Ridge Regression}
With regression on certain datasets (for example, one with many features but not enough data), we often run into the issue of overfitting. In other words, our model may fit the training set very well but fails to generalize. We can address this using regularization, which includes the squared norm of the weight vector as a penalty, scaled by a factor $\lambda$ that induces bias to counter variance.

\subsection{Implementation and Testing}
Using the closed form $\theta = (\phi ^ {T} \phi + \lambda I)^{-1} \phi ^ T Y$, we implemented ridge regression in Python, using the numpy library. Then, using the simple data from Bishop's Figure 1.4, we ran ridge regression using $M$ ranging from $0$ to $10$ and $\lambda$ from $0, 0.0001, 0.0002 \dots, 5.24288$. We noted that for larger values of $M$, while the model fit the data well, it did not generalize well past the given range of values, suggesting that small to intermediate $M$ values may fit the data better. Furthermore, for larger values of $M$, changing $\lambda$ slightly had a greater effect as the elements of $\theta$ were fairly large, so penalizing the weight vector incurred a large cost. Using $\lambda = 0$ as a baseline, we were able to witness the effects of regularization in smoothing out curves. Considering $\lambda$ in particular, we note that large lambda values result in high bias and thus underfitting, while low lambda values (approximately 0) result in high variance and thus overfitting, in line with our previous assumptions.

\begin{figure}
\label{p3figure1}
\includegraphics[scale=0.4]{bishop7.png}
\caption{Ridge regression with $M = 7, \lambda = 0.00128$}
\end{figure}

\subsection{Model Selection}
Regarding the process of model selection, we initially began with $M$ again ranging from $0$ to $10$ and 200 values for $\lambda$ ranging from $10 ^ (-8)$ to $1.726$, generated by continually multiplying $1.1$ to an initial $\lambda$. Then, using the training data set, we were able to generate $\theta$ values for all combinations of $M$ and $\lambda$. We then tested each of these $\theta$ values on the validation set, and computed the error generated. Then, using the models with the smallest error produced, we applied them on the testing set to get our final results. \\ \\ 
From testing our $\theta$'s on the validation set, we were able to secure $M = 4, \lambda = 0.65536; M = 3, \lambda = 0;$ and $M = 3, \lambda = 10 ^ {-5}$ with respective errors $3.208586, 3.6360459, 3.636073$. Considering the testing set, we arrived at $M = 3, \lambda = 10 ^ {-5}$ as our model, with the lowest error of $0.2496965$. Graphs have been produced below.

\subsection{BlogFeedback}
Ridge regression is especially suitable here as the blog dataset is one with many features (280) in fact, but not too much data (most of the identified values being $0$).This was slightly different from our previous model selection process as we are using a simple linear model here. Thus, as the focus was on finding the appropriate value for $\lambda$, we used an exhaustive approach and considered values in range $10 ^ {-5}$ to $1$ in increments of $10 ^ (-5)$. Proceeding through identifying $\theta$ based on the training dat set, validating the model with the validation dataset, and finally testing with the testing dataset, we arrived at $\lambda = 0.0091315$, with a surprisingly high error of $8051727.304087$.
